%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Preamble
\documentclass{article}
\usepackage{amsmath,amssymb,amsthm,fullpage}
\usepackage[a4paper,bindingoffset=0in,left=1in,right=1in,top=1in,
bottom=1in,footskip=0in]{geometry}
\newtheorem*{prop}{Proposition}
%\newcounter{Examplecount}
%\setcounter{Examplecount}{0}
\newenvironment{discussion}{\noindent Discussion.}{}
\pagenumbering{gobble}
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Heading
%\title{CS21 Midterm}
%\author{Matt Lim}
%\date{February 12, 2014}
%\maketitle



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Problem 1
\section*{Problem 1, CS21 Set 5, Matt Lim}
\begin{description}
    \item[(a)] False. To prove this we must find a language L that has the
        following properties: L $\in$ EXP and L $\notin$ TIME$(2^{n^{100}}$).
        To find this L, we will use the Time Hierarchy Theorem. Consider
        the proper complexity function $f(n) = 2^{n^{100}} \ge n$. Then we
        have that $f(2n)^{3} = (2^{2n^{100}})^{3}$. So we have that, by
        the Time Hierarchy Theorem,
        TIME($2^{n^{100}}$) $\subsetneq$ TIME($(2^{2n^{100}})^{3}$) $\subseteq$
        EXP. This means that we can choose an L that is in
        TIME($(2^{2n^{100}})^{3}$) (and thus in EXP) that is not in
        TIME($2^{n^{100}}$). This proves (a) to be false.

    \item[(b)] True. This is because we showed in lecture that P $\subsetneq$
        EXP, which means that either P $\neq$ NP or NP $\neq$ EXP (or both).
    \item[(c)] False. Since, due to the problem, we can assume that an
        NP-complete language has an $O(n^{k})$-time algorithm for some fixed
        $k$, we have that NP = P. So then we can do the following.
        To prove this we must find a language L that is in P
        but not in TIME($n^{k}$). To find this L, we will use the Time Hierarchy
        Theorem. Consider the proper complexity function $f(n) = n^{k}$. Then
        we have that $f(2n)^{3} = (2n)^{3k}$. So we have that, by the Time
        Hierarchy Theorem, TIME($n^{k}$) $\subsetneq$ TIME($(2n)^{3k}$)
        $\subseteq$ P. This means that we can choose an L that is in
        TIME($(2n)^{3k}$) (and thus in P) that is not in TIME($n^{k}$). This
        proves (c) to be false.
    \item[(d)] False. Consider the language L = $\Sigma ^{*}$. This language is
        clearly in P (everything is accepted) and thus in NP. To be NP-complete,
        every language A in NP
        must reduce to L. But we cannot reduce any language A in NP to L because
        it is impossible to map no to no, since L does not have any instances of
        no to map to. Thus L is not NP-complete, yet is in NP. So (d) is false.
    \item[(e)] True. This is because NP $\subseteq$ EXP. So since every
        EXP-complete language has the characteristic that every language
        A $\in$ EXP reduces to it, and since we have that NP $\subseteq$ EXP and
        thus that every language in NP is in EXP, we can conclude that every
        language in NP reduces to every EXP-complete language.
\end{description}
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Problem 2
\section*{Problem 2, CS21 Set 5, Matt Lim}
To show
\[ \text{3-COLORABLE} = \{G : G \text{ is 3-colorable}\} \]
is NP-complete, we must show that it is in NP and that all NP-problems are
polynomial time reducible to it. The first part is easy; a certificate is simply
a 3-colorable graph. To prove the second part, we show that 3-SAT is polynomial
time reducible to 3-COLORABLE. The following details the reduction.

Our reduction function $f$ will map a 3-CNF formula $\varphi$ to a graph $G$.
The graph will look as follows:
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Problem 3
\section*{Problem 3, CS21 Set 5, Matt Lim}
To show that (3,3)-SAT is NP-complete, we must show that it is in NP and that
all NP-problems are polynomial time reducible to it. The first part is easy;
a certificate is simply a CNF formula with at most 3 literals per clause, and
at most 3 occurrences of any variable. To prove the second part, we show that
3-SAT is polynomial time reducible to (3,3)-SAT. The following details the
reduction.

Our reduction function $f$ will map a 3-CNF formula $\varphi$ to a new 3-CNF
formula $\varphi_{2}$. If $\varphi$ itself has at most 3 occurrences of any
variable, then we can just make $\varphi_{2}$ a copy of $\varphi$. Else, if
$\varphi$ has more than 3 occurrences of any variable, we can do the following
thing to each of these variables.

Let an arbitrary variable $r$ repeat $n > 3$ times. Then we will introduce
$n$ new variables $y_{1}, y_{2}, \dots , y_{n}$. We want all the $y_{i}$s to
be equivalent to each. That is, we want $y_{1} \Leftrightarrow y_{2}
\Leftrightarrow \dots \Leftrightarrow y_{n}$. To get this, it suffices to show
that $y_{1} \Rightarrow y_{2} \Rightarrow \dots \Rightarrow y_{n} \Rightarrow
y_{1}$. To do this in CNF format, it will look like the following:
\[ (y_{2} \vee \overline{y_{1}}) \wedge (y_{3} \vee \overline{y_{2}})
\wedge \dots (y_{1} \vee \overline{y_{n}}) \]
Then, we will add the above clauses to $\varphi$ (separated by $\wedge$s), and
replace each $i$th instance of $r$ with $y_{i}$.

Now we will show that this reduction function $f$ maps yes to yes and no
to no. First we will show it maps yes to yes. So, given that $\varphi$ has a
satisfying assignment, we want to show that $\varphi_{2}$ is satisfiable
and that it is in (3,3)-SAT. To do this, we will consider our reduction
function. Note that for any arbitrary variable $r$ repeated $n > 3$ times, our
reduction function adds new variables $y_{1}, y_{2}, \dots , y_{n}$, each of
which appear twice (once in the added clauses, once in the replacements).
Note that the complement of each $y_{i}$ appears once. Now
we must consider what happens if $\neg r$ repeats $k > 3$ times as well. That
is, both $r$ and $\neg r$ repeat more than three times. If this is the case,
then we can first do the replacements for whichever variable repeats more.
So, WLOG, consider the case where $r$ repeats more. Then we will have
$y_{1} \Leftrightarrow y_{2} \Leftrightarrow \dots \Leftrightarrow y_{n}$.
Since we have this, we also have that $\overline{y_{1}} \Leftrightarrow
\overline{y_{2}} \Leftrightarrow \dots \Leftrightarrow \overline{y_{n}}$. Thus,
to replace our $\overline{r}$s, we can simply replace each $i$th instance
of $\overline{r}$ with $\overline{y_{i}}$. Then we have that we have added
an instance of each $\overline{y_{i}}$, bringing the total number of each one
up to two. So, we have that our reduction function gives us a $\varphi_{2}$
that has at most 3 literals per clause and at most 3 occurrences of each
variable. Now we must show that it is satisfiable to complete the proof of
yes maps to yes. To prove this, we must only show that our added clauses
are satisfiable, since our replacement method does not change the
satisfiability of the original clauses (since all the replacements for a
given repeated variable are equivalent). This is clearly true, since we can
just either make all the $y_{i}$'s to be true or all of them to be false.
So we have that yes maps to yes.

Now we will show that this reduction function $f$ maps no to no. To prove this,
consider again the fact that our replacement method does not change the
satisfiability of the original clauses (since all the replacements for a given
repeated variable are equivalent). Thus, the non-added-in clauses in
$\varphi_{2}$ are unsatisfiable. And adding in the extra clauses can do nothing
to change this, since all the clauses are separated by $\wedge$s. So we have
that no maps to no.

Now we must only prove that our reduction function $f$ is polynomial time
computable. A variable can repeat no more than $3n$ times, where $n$ is the number
of clauses. And there can be no more than $3n$ repeated variables. This means
that for each repeated variable, we do no more than $3n$ replacements, and
add no more than $3n + 1$ clauses. Clearly the function is polynomial.
So we have that our reduction function is polynomial.
\newpage
\end{document}



